---
title: "HARtools: Basic Use"
author: "John D Harrison"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{HARtools: Basic Use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 80)
```
The goal of this vignette is to describe the basic functions in the 
`HARtools` package.

## HAR format

HAR stands for HTTP Archive. This is a common format for recording HTTP 
tracing information. This file contains a variety of information. It has a 
record of each object being loaded by a browser. Each of these objects’ 
timings is recorded. The common extension for these files is .har.  The 
specification for this format is produced by the Web Performance Working
Group of the World Wide Web Consortium (W3C).

An outline of the specification can be viewed at [softwareishard.com](http://www.softwareishard.com/blog/har-12-spec/)

## HARtools basics

`HARtools` is a lightweight package containing a number of tools to
read/parse, visualise and write files in the HAR format.

### Reading HAR files
`readHAR` is the function whoose aim is to read read/parse HAR objects. It
currently can read/parse a number of formats. It takes an argument `har` 
which currently can be a character string, an R list, a URL or a file
containing HAR data. It also carries out basic validation on the HAR 
content (checking mandatory/optional fields).

#### Reading/Parsing from file

The `HARtools` package contains a number of examples of HAR files. We will
use one of these (HAR data from browsing `google.com`):

```{r}
library(HARtools)
har <- readHAR(system.file(package = "HARtools", 
                           "exdata", "google.com.har"))
har
```
```{r}
names(har)
names(har[["log"]])
```

So we see that a HAR object is a list with root `log` and children as 
defined per the specification. 

#### Reading/Parsing a URL
The same HAR data we sourced from file is available [online](http://www.janodvarko.cz/har/viewer/examples/google.com.har).

```{r}
harData <- "http://www.janodvarko.cz/har/viewer/examples/google.com.har"
harURL <- readHAR(harData)
identical(harURL, har)
```

#### Reading/Parsing an R character string

`readHAR` can read/parse an R character string

```{r}
fileName <- system.file(package = "HARtools", "exdata", "google.com.har")
harString <- readChar(fileName, file.info(fileName)$size)
print(substr(harString, 1, 5*80))
harSTR <- readHAR(harString)
identical(harSTR, har)

```

#### Reading/Parsing an R list

`readHAR` can read/parse an R list.

```{r}
harL <- jsonlite::fromJSON(system.file(package = "HARtools",
                                          "exdata", "google.com.har"),
                              simplifyVector = FALSE)
harLIST <- readHAR(harL)
identical(harLIST, har)

```

### Writing HAR files

Parsed HAR objects of class `har` can be written to file using the
`writeHAR` function:

```{r}
harFile <- tempfile(fileext = ".har")
writeHAR(har, harFile)
har2 <- readHAR(harFile)
identical(har, har2)
```

`writeHAR` exports the HAR in JSON format.


## Visualising HAR data

### HARviewer

`HARtools` includes a `HARviewer` function. `HARviewer` is a `htmlwidget`
which utilises the [PerfCascade](https://github.com/micmro/PerfCascade) to 
create a waterfall view of an HAR object.

```{r}
har <- readHAR(system.file(package = "HARtools", "exdata",
                           "r-project.org.161028_W2_11MA.har"))
hv <- HARviewer(har, width="100%", height="100%")
# view in R
hv

```

The above HAR data was obtained from testing the [r-project](http://r-project.org)
webpage using [WebPageTest](https://www.webpagetest.org/). 

### Embedding HAR charts in a shiny app

`HARtools` has a `renderHARviewer` function and a corresponding `HARviewerOutput`
function which can be used in a `shiny` server and ui respectively to 
embed a `HARviewer` waterfall chart in a shiny app:

```{r, eval=FALSE}
library(shiny)
if(interactive()){
  library(HARtools)
  har <- readHAR(system.file(package = "HARtools", "exdata",
                             "r-project.org.161028_W2_11MA.har"))
  hv <- HARviewer(har)
  
  app <- shinyApp(
    ui = fluidPage(
      HARviewerOutput("myHAR")
    ),
    server = function(input, output) {
      output$myHAR <- renderHARviewer(hv)
    }
  )  
  runApp(app)
}
```

## Obtaining HAR data

### Using a browser
Most modern browser have developer tools that often have the option to 
export a browsing session as HAR data see for example this 
[article](https://confluence.atlassian.com/kb/generating-har-files-and-analysing-web-requests-720420612.html).

### Using Browser Mob Proxy

[Browser Mob Proxy](https://bmp.lightbody.net/) captures performance data 
for web apps (via the HAR format), as well as manipulate browser behavior 
and traffic, such as whitelisting and blacklisting content, simulating 
network traffic and latency, and rewriting HTTP requests and responses.

An R client for BMP is available at 
https://github.com/johndharrison/bmproxy.

### Using WebPageTest

[WebPagetest](https://github.com/WPO-Foundation/webpagetest) is an open 
source project that is primarily being developed and supported by Google 
as part of their efforts to make the web faster.

WebPagetest is a tool that was originally developed by AOL for use 
internally and was open-sourced in 2008 under a BSD license. The platform 
is under active development on GitHub and is also packaged up periodically
and available for download if you would like to run your own instance.

The online version at [www.webpagetest.org](https://www.webpagetest.org) 
is run for the benefit of the performance community with several companies 
and individuals providing the testing infrastructure around the globe.

An R client for WPT is available at 
https://github.com/johndharrison/webpagetestr. It allows the user to send
test requests to [www.webpagetest.org](https://www.webpagetest.org) or any
private instance of WPT a user may have access to.

